Double, Double, Spin plus Nubble: Two Stars and Two Signals in PTFO 8-8695

### 2020-04-27 paper/ms.tex
Update ms.tex

\begin{abstract}
  PTFO$\,$8-8695 is a star in the 7--10 million year
  old Orion-OB1a cluster that shows brightness dips that resemble planetary transits.
  While strong evidence against the planet hypothesis has already been provided,
  the possibility remains debated in the literature.
  To obtain further clues, we inspected data from the
  NASA {\it Transiting Exoplanet Survey Satellite} (TESS) and the ESA Gaia mission.
  The Gaia data show that
  PTFO$\,$8-8695 is a photometric binary with respect to members of
  its kinematic group, and independently suggest that it is an
  astrometric binary.  The TESS lightcurve shows two different photometric
  periods. The variability is dominated by a sinusoidal signal with a period
  of 11.96\,$hr, presumably caused by stellar rotation.  Also
  present is a 10.74-day signal consisting of a not-quite sinusoid interrupted
  by 1.2\% hour-long dips, the type of signal previously
  intepreted as planetary transits.
  The phase of the dips is nearly 180$^\circ$ away from the phase of the originally reported dips,
  which, as noted previously, makes them difficult to explain as planetary transits.
  Instead, we believe that PTFO$\,$8-8695 is a pair of young
  and rapidly rotating M dwarfs, one of which shows the same ``transient-dipper''
  behavior that been seen in at least 5 other cases.
  The origin of these transient dips is still unknown but likely involves circumstellar material.
\end{abstract}

@@

If \ptfob\ were a planet, it would be exceptional.  At below 10
million years of age, it could be the youngest hot Jupiter known
\citep{van_eyken_ptf_2012}.  Its orbital period of only 10.7 hours
around a weak-lined T Tauri M dwarf would also give it the shortest
period of any known hot Jupiter.  With such a short period, it would
almost certainly have filled its Roche lobe, and would be actively
losing mass to its host star.  Not only that, but the rapidly rotating
host star would also be oblate and gravity darkened, and so the
planet's orbit would likely precess into and out of transitability
\citep{barnes_measurement_2013,ciardi_followup_2015,kamiaka_revisiting_2015}.

Other lines of evidence would imply further planetary ``firsts'' for
this planet candidate.  One first would be that its transits are about
three times deeper in optical bandpasses ({\it e.g.,} $g$-band) than
in the near-infrared ({\it e.g.}, $z$-band)
\citep{onitsuka_multicolor_2017,tanimoto_evidence_2020}.  A cloud-free
hydrogen-dominated planetary atmosphere cannot explain such a
wavelength dependence.  The planet might therefore be surrounded by a
dust cloud \citep{tanimoto_evidence_2020}.

Another first could be the direct detection of H$\alpha$ emission from
the planet itself \citep{johnskrull_h_2016}.  While the stellar
chromosphere emits in H$\alpha$, it seems that there is an additional
excess H$\alpha$ emission that could be in phase with the planetary
orbit.  The average velocity width of the excess H$\alpha$ emission is
87$\,$km$\,$s$^{-1}$, and its equivalent width is 70-80\% that of the
stellar chromosphere \citep{johnskrull_h_2016}.  The proposed
explanation is that outflowing mass from the planet could explain this
excess emission as well \citep{johnskrull_h_2016}.

There are perhaps a few challenges to the planetary interpretation (if
these ``features'' are not already seen as such).  They include that
the planet does not seem to emit infrared radiation in occulation, at
least at the expected amplitude \citep{yu_tests_2015}.  In addition,
despite measurement attempts by multiple investigators, \ptfob\ does
not seem to show the Rossiter effect at the amplitude expected given
the rapid stellar rotation and large planet size
\citep{yu_tests_2015,ciardi_followup_2015}.  Finally, detailed
modelling of the ``precession + gravity darkening'' transits has shown
that the necessary degree of gravity darkening is too great, given the
spectroscopically observed equatorial velocity
\citep{howarth_reappraisal_2016}.  Additionally, as the
gravity-darkened star precessed about its rotation axis, it would need
to show photometric variability that has not been observed.

While the planetary interpetation clearly faces challenges,
alternative explanations do as well.  High-latitude accretion hotspots
might produce the observed H$\alpha$ variability, but require
fine-tuning to produce dips of the appropriate duration. Furthermore,
\ptfo\ does not have an infrared (IR) excess associated with the
presence of dust at the inner edge of a primordial disk \citep[{\it
e.g.},][Figure~18]{yu_tests_2015}.  Low-latitude starspots, hot or
cold, struggle to produce the necessary dip durations.

One potentially relevant fact is that between 0.1\% and 1\% of rapidly
rotating low-mass stars in $\mathcal{O}$(10)$\,$Myr old associations
show narrow dips in phase with strong stellar rotation signals
\citep{rebull_usco_2018}.  The dips can persist over months, but their
depths often vary, and sometimes change immediately after stellar
flares.  The explanation proposed by \citet{stauffer_orbiting_2017}
and \citet{david_transient_2017} to explain this novel class of
variable stars is that a circumstellar cloud of gas might be orbiting
near the Keplerian co-rotation radius.  To date, this explanation has
not clearly been viable for \ptfo, because a clear determination of
the stellar rotation period has not yet been possible
\citep{van_eyken_ptf_2012,koen_multicolour_2015,raetz_yeti_2016}.

We begin in Section~\ref{sec:observations} by describing newly
available observations from TESS \citep{ricker_transiting_2015} and
Gaia \citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows
two distinct signals, which we extract and analyze in
Section~\ref{sec:tess}.  A long-period sinusoid repeats every 12.0
hours, and is probably stellar rotation.  A short-period dip
superposed on an asymmetric sinusoid repeats every 10.7 hours.
Analyzing the Gaia data in Section~\ref{sec:gaia}, we show that
relative to its kinematic group, \ptfo\ is a photometric binary, and
suggest it also an astrometric binary.  We discuss the puzzle pieces
in Section~\ref{sec:discussion}.  The orbital phase of the dip seems
to have instantaneously changed over the past decade.  In addition,
other young stars show lightcurve morphologies similar to the
short-period signal.  We therefore argue that PTFO$\,$8-8695 is a
binary M dwarf in which one star shows a rotation signal, and the
other shows transient dipping potentially caused by eclipses of
plasma at the Keplerian co-rotation radius.
Section~\ref{sec:conclusions} summarizes our main points.

@@

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 19).}
    {\it Top}: ``Raw'' \texttt{PDCSAP} mean-subtracted relative flux
    versus time. The beat period of 4.48 days is visible by eye.  The
    preferred model plotted underneath the data includes 2 harmonics
    at the long period $P_{\rm \ell}$, plus 2 harmonics and a transit
    at the short period $P_{\rm s}$.
		{\it Upper middle}: Long-period signal, equal to the raw signal
		minus the short-period signal.
		{\it Lower middle}: Short-period signal, equal to the raw signal
		minus the long-period signal.
		{\it Bottom}: residual.  The data are binned from 2 to 10 minute
		cadence as a convenience for plotting and fitting.
		\label{fig:splitsignal}
	}
\end{figure*}

\begin{figure*}[hbtp]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 20).}
		Panels are as in Figure~\ref{fig:splitsignal}.
		\label{fig:splitsignalii}
	}
\end{figure*}

@@

\ptfo\ was observed by TESS with Camera 1, CCD 1, from December 15,
2018 to January 6, 2019, during the sixth sector of science operations
\citep{ricker_transiting_2015}.  The star was designated TIC 264461976
in the TESS Input Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.
The pixel data for an $11\times11$ array surrounding \ptfo\ were
averaged into 2-minute stacks by the onboard computer.  Each
2048$\times$2048 image from the CCD was also averaged into 30-minute
stacks, and saved as a ``full frame image'' (FFI).

The 2-minute stacks for \ptfo\ were reduced to lightcurves by the
Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.  Our main analysis used the resulting
Presearch Data Conditioning (PDC) lightcurve.  The PDC lightcurve
aperture used pixels chosen to maximize the SNR of the total flux of
the target \citep{smith_kepler_apertures_2017}.  Non-astrophysical
variability was removed by fitting out trends common to many stars
\citep{smith_kepler_PDC_2017}.

As an independent check on the shorter cadence SPOC light-curve, we
separately processed the 30-minute image stacks as part of the Cluster
Difference Imaging Photometric Survey (CDIPS;
\citealt{bouma_cluster_2019}).  Our CDIPS lightcurve  of choice used a
circular aperture with radius 1 pixel.

To clean the data, we removed all points with non-zero quality flags
\citep[{\it e.g.},][]{tess_data_product_description_2018}.  We also
masked out the first and last 6 hours of each orbit, since there is
often systematic red noise during those times.  Both the CDIPS and PDC
lightcurves showed a clear discontinuous ``jump'' in the last few days
of orbit 20, which seemed likely to be an instrumental systematic.  We
correspondingly masked out times from BJD 2458488.3 until the end of
the orbit.  The PDC lightcurve initially had 15{,}678 points.  The
quality-flag cut removed 854 points; masking the orbit edges removed an
additional 716; removing the final few days of orbit 20 removed an
additional 1079.  After cleaning, 83\% of the initial flux
measurements remained.

We normalized these points by dividing out the median flux. We opted
to then subtract by unity to simplify subsequent interpretation.  Many
of these and subsequent processing steps were performed using
\texttt{astrobase}~\citep{bhatti_astrobase_2018}.

@@

\subsubsection{Astrometry of \ptfo}

Between July 25, 2014 and May 23, 2016, Gaia measured about 300
billion centroid positions of 1{.}6 billion stars
\citep{gaia_collaboration_gaia_2016,lindegren_gaiasoln_2018,gaia_collaboration_gaia_2018}.
In the Gaia second data release (DR2), these CCD observations were
used to estimate positions, proper motions, and parallax for the
brighest 1{.}3 billion stars, including \ptfo\
\citep{lindegren_gaiasoln_2018}.  There were 121 ``good'' observations
of \ptfo, that is observations that were not strongly down-weighted in
its astrometric solution.  \ptfo\ was assigned a Gaia DR2 identifier
of 3222255959210123904.  It photometric brightness was measured using
selected bands ($G$, $Rp$, and $Bp$) of the Gaia Radial Velocity
Spectrometer \citep{cropper_gaia_2018,evans_gaia_2018}.  We accessed
the pipeline parameters for \ptfo\ using the Gaia
archive\footnote{\url{gea.esac.esa.int/archive/}}.

The majority of Gaia's derived parameters for \ptfo\ agreed with
expectations from former studies
\citep{briceno_cida_2005,van_eyken_ptf_2012}.  One novelty however was
that Gaia DR2 detected a 10.3$\sigma$ ``astrometric excess''.
We comment on the significance and
interpretation of this excess in Section~\ref{sec:gaia}.

@@

Gaia also provided astrometric parameters for tens of thousands of
young stars in the Orion complex.  Stellar populations in giant
molecular cloud complexes are not monolithic; substructured groups are
the norm \citep{briceno_lowmassOB_2007}.  The Orion molecular cloud
complex in particular has numerous subgroups, with ages spanning 0.5
to 15$\,$Myr. For an incomplete sampling, see for instance
\citet{briceno_cida_2005,jeffries_kinematic_2006,briceno_25_2007,kounkel_apogee2_2018}
and \citet{briceno_cidaII_2019}.

\ptfo\ was initially identified as a member of the Orion$\,$OB1a
sub-association by \citet{briceno_cida_2005} through combined
photometry and spectroscopy.  Later work by \citet{briceno_25_2007}
clarified that \ptfo\ was in a kinematically distinct subgroup of
Orion~OB1a, named the ``25$\,$Ori'' group after its brightest
member. The 25$\,$Ori group had an isochrone age of 7--10$\,$Myr, and
had a smaller disk fraction than younger nearby sub-associations
\citep{hernandez_spitzer_ob1_2007}.

With the Gaia astrometry, it has become clear that 25$\,$Ori itself
has further subgroups
\citep{kounkel_apogee2_2018,briceno_cidaII_2019}.  In describing the
cluster membership of \ptfo, we follow the notation and results of
\citet{kounkel_apogee2_2018}.  These authors combined astrometric data
from Gaia DR2 with near-infrared spectra from APOGEE-2
\citep{gunn_sdss_2006,majewski_apache_2017,blanton_sloan_2017,zasowski_target_2017,cottle_apogee2_2018}.
They performed a hierarchical clustering on the six dimensional
position and velocity information to identify subgroups within the
Orion complex.  From smallest to largest, \ptfo\ was identified
as a member of the following hierarchical subgroups:
\begin{equation}
  %{\rm PTFO\,8\text{-}8695}
  % \in
  {\rm 25\,Ori\text{-}1}
  \subset {\rm 25\,Ori}
  \subset {\rm Orion\ OB1a}
  \subset {\rm Orion\ D},
\end{equation}
where from set-notation, `$\subset$' denotes ``is a proper subset
of''.  25$\,$Ori-1 is the largest subgroup of 25$\,$Ori, with 149
identified members.  Its mean age from its CMD was found to be
6.9$\,$Myr, and from its HR diagram 8.5$\,$Myr
\citep{kounkel_apogee2_2018}.  \citet{kounkel_apogee2_2018} identified
seven other smaller groups in the Orion complex near the Be star
25$\,$Ori. These groups received higher numbers, {\it e.g.},
25$\,$Ori-2 (${\rm Age}_{\rm CMD} =15.1\,$Myr; ${\rm Age}_{\rm CMD}
=12.9\,$Myr; see also \citealt{briceno_cidaII_2019}).

These details concerning the group membership for one object may seem
cumbersome to those accustomed to comparing ``young cluster members''
with ``old field stars''.  Though all members of the Orion complex are
indeed young relative to the field, these details are essential for
assessing any evidence for photometric binarity in \ptfo, because of
the degeneracy between stellar luminosity and age for
pre-main-sequence stars.  Having a clean sample of tightly spatially
and kinematically associated reference stars minimizes contamination
not just from the field, but from older and younger members of the
Orion complex itself.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f3.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Phase-folded long and short-period signals.}
    {\it Top}: Long-period signal, as in Figure~\ref{fig:splitsignal}.
    {\it Bottom}: Short-period signal. The reference phase is set to
    the dip.  Gray points are the 10 minute cadence
    \texttt{PDCSAP} flux.  Black points are binned to 100 points per
    period.
		\label{fig:phasefold}
	}
\end{figure*}

@@

Our initial inspection of the TESS lightcurve, in both its 2-minute
PDCSAP and 30-minute FFI forms, showed a strong sinusoidal beat signal
(Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}, top
panel).

As a precursor to more detailed analysis, we calculated generalized
Lomb-Scargle periodograms using \texttt{astrobase}
\citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}.
The two largest peaks in the Lomb-Scargle periodogram of the
lightcurve were separated at a ``short'' period $P_{\rm s}
\approx 0.448\,{\rm days}$ and a ``long'' period $P_{\rm \ell} \approx
0.499\,{\rm days}$.  The $P_{\rm \ell}$ peak had the greater power of
the two.  Smaller harmonics surrounding each of these two dominants
peaks were also present.

@@

The peak-to-peak lightcurve amplitude at maximum, when the two signals
constructively interfere, is about 14\%.  At minimum, the peak-to-peak
amplitude is about 6\%.  Assuming the signals are just two sinusoids,
algebra tells us that the peak-to-peak amplitudes should therefore be
10\% for the long-period signal, and 4\% for the short-period signal.
These order-of-magnitude numbers will turn out to be roughly correct.

Initial signal-processing experiments fitting out splines or sinusoids
showed that after subtracting out the long-period signal, the
short-period signal dominated the periodogram, and vice-versa.
However it quickly became clear that it would be beneficial to
simultaneously model each signal, in order to preserve the
power at each frequency.

@@

We considered nine models, with the number of harmonics per frequency
$N$ and $M$ ranging from one to three.  To select our preferred model,
we used the Bayesian information criterion (Table~1).  The model with
the lowest BIC had two harmonics at the short 10.74$\,$hr period, and
two harmonics at the long 11.96$\,$hr period.  The next-best model had
an additional harmonic at the longer period ($M=3$), but was otherwise
identical.  All nine models have reduced $\chi^2$ ranging between 1.37
and 1.51, which suggests a plausible though imperfect agreement
between the data and fitting results.  The best-fit parameters for the
lowest BIC model are given in Table~2.

@@

\subsection{Long Period Sinusoid}

The standard interpretation for 11.96$\,$hour sinusoidal modulations
of a pre-main-sequence M dwarf is stellar rotation.  This is the
dominant signal in the system with 10\% amplitude, and there is no
evidence to suggest that this signal has any other origin.

>>>>

\begin{abstract}
  PTFO$\,$8-8695 (CVSO\,30) is a star in the 7--10 million year
  old Orion-OB1a cluster that shows brightness dips that resemble planetary transits.
  While strong evidence against the planet hypothesis has already been provided,
  the possibility remains debated in the literature.
  To obtain further clues, we inspected data from the
  NASA {\it Transiting Exoplanet Survey Satellite} (TESS) and the ESA Gaia mission.
  The Gaia data show that
  PTFO$\,$8-8695 is a photometric binary with respect to members of
  its kinematic group, and independently suggest that it is an
  astrometric binary.  The TESS lightcurve shows two different photometric
  periods. The variability is dominated by a sinusoidal signal with a period
  of 11.96\,$hr, presumably caused by stellar rotation.  Also
  present is a 10.74-day signal consisting of a not-quite sinusoid interrupted
  by 1.2\% hour-long dips, the type of signal previously
  intepreted as planetary transits.
  The phase of the dips is nearly 180$^\circ$ away from the phase of the originally reported dips,
  which, as noted previously, makes them difficult to explain as planetary transits.
  Instead, we believe that PTFO$\,$8-8695 is a pair of young
  and rapidly rotating M dwarfs, one of which shows the same ``transient-dipper''
  behavior that been seen in at least 5 other cases.
  The origin of these transient dips is still unknown but likely involves circumstellar material.
\end{abstract}

@@

We wish \ptfob\ were a planet. It would be quite exceptional. It would
be the youngest known hot Jupiter
\citep{van_eyken_ptf_2012}, orbiting a T Tauri star in the Orion-OB1a cluster.
It would have the shortest orbital period (10.7 hours) of any hot Jupiter.
With such a short period, it would
probably be filling its Roche lobe, and actively
losing mass to its host star.  Not only that, but the rapidly-rotating
host star is probably oblate enough to torque the planet's orbit
into and out of the transiting configuration on a time scale of years
\citep{barnes_measurement_2013,ciardi_followup_2015,kamiaka_revisiting_2015}.

Another first would be the direct detection of H$\alpha$ emission from
the planet itself \citep{johnskrull_h_2016}.  In addition to the
chromospheric H$\alpha$ emisson, it seems that there is an additional
H$\alpha$ emission with radial velocity variations in phase with the planetary
orbit.  The average velocity width of the excess H$\alpha$ emission is
87$\,$km$\,$s$^{-1}$, and its equivalent width is 70-80\% that of the
stellar chromosphere \citep{johnskrull_h_2016}.  The proposed
explanation is that the emission is from hot material flowing away
from the planet \citep{johnskrull_h_2016}.

However, the observed signals have some peculiarities that
make the planet seem even more unusual, to the point that they cast into doubt the premise
that \ptfob\ is real.
First, the transit-like brightness dips are about
three times deeper in optical bandpasses ({\it e.g.,} $g$-band) than
in the near-infrared ({\it e.g.}, $z$-band)
\citep{onitsuka_multicolor_2017,tanimoto_evidence_2020}.  An ordinary
atmosphere expected for a Jovian planet would not lead to such a strong
color-dependence of the transits.  
Second, the planet does not seem to emit as much infrared radiation
as would be expected for such a hot Jovian planet \citep{yu_tests_2015}.
Third, despite measurement attempts by multiple investigators, \ptfob\ does
not seem to show the Rossiter effect at the amplitude expected given
the rapid stellar rotation and large planet size
\citep{yu_tests_2015,ciardi_followup_2015}.
Fourth, the phase of the dips within the overall period of photometric variability
has changed drastically over the years since their initial discovery.
To counter these objections, it has been proposed 
that the planet may be much smaller than Jupiter and that the dips
are produced by dust clouds emitted from the planet \citep{tanimoto_evidence_2020}.

A separate issue is that the brightness dips
change shape over many orbital cycles. This was initially explained by
\cite{barnes_measurement_2013} as the natural effects of gravity darkening
However, \cite{howarth_reappraisal_2016} argued
that the necessary amplitude of gravity darkening is too large to be realistic, given the
spectroscopically-determined rotation velocity.
Additionally, as the
gravity-darkened star precessed about its rotation axis, it would 
show photometric variability that has not been observed.

While the planetary interpetation clearly faces challenges,
there is no completely satisfactory alternate explanation.  High-latitude accretion hotspots
might produce the observed H$\alpha$ variability, but require
fine-tuning to produce dips of the appropriate duration. Furthermore,
\ptfo\ does not have an infrared (IR) excess associated with the
presence of dust at the inner edge of a primordial disk \citep[{\it
e.g.},][Figure~18]{yu_tests_2015}.  Low-latitude starspots, hot or
cold, struggle to produce photometric features as short as some
of the observed dips.

A relevant fact is that between 0.1\% and 1\% of rapidly
rotating low-mass stars in $\mathcal{O}$(10)$\,$Myr old associations
show short-duration dips as part of their overall periodic variability
\citep{rebull_usco_2018}.  The dips can persist over months, but their
depths often vary, and sometimes change immediately after stellar
flares.  The explanation proposed by \citet{stauffer_orbiting_2017}
and \citet{david_transient_2017} to explain this novel class of
variable stars is that a circumstellar cloud of gas is orbiting
near the co-rotation radius.  To this point, though, it has not
been clear if this explanation applies to
\ptfo, because the determination of
the stellar rotation period has been somewhat ambiguous
\citep{van_eyken_ptf_2012,koen_multicolour_2015,raetz_yeti_2016}.

We begin in Section~\ref{sec:observations} by describing newly
available observations from TESS \citep{ricker_transiting_2015} and
Gaia \citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows
two different periodic signals, which we analyze in
Section~\ref{sec:tess}.
The Gaia data, analyzed in Section~\ref{sec:gaia}, show that
\ptfo\ is a photometric binary, and suggest it also an astrometric binary.
We discuss the pieces of the puzzle
in Section~\ref{sec:discussion}, and summarize the situation in
Section~\ref{sec:conclusions}.
% JNW: this kind of stuff is for the abstract/conclusion, not intro.
%The phase of the dip within the seems
%to have instantaneously changed over the past decade.  In addition,
%other young stars show lightcurve morphologies similar to the
%short-period signal.  We therefore argue that PTFO$\,$8-8695 is a
%binary M dwarf in which one star shows a rotation signal, and the
%other shows transient dipping potentially caused by eclipses of
%lasma at the Keplerian co-rotation radius.

@@

% JNW: How about more descriptive labels? f --> Total flux
% fl = Longer-period signal
% fs = Shorter-period signal
% last panel, Residuals

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 19).}
    {\it Top}: The ``raw'' (\texttt{PDCSAP} mean-subtracted) relative flux
    data. The beat period of 4.48 days is visible by eye.  The
    blue curve is a model including 2 harmonics
    at the longer period $P_{\rm \ell}$, plus 2 harmonics and a transit
    at the shorter period $P_{\rm s}$.
		{\it Upper middle}: Longer-period signal, equal to the raw signal
		minus the shorter-period signal.
		{\it Lower middle}: Shorter-period signal, equal to the raw signal
		minus the longer-period signal.
		{\it Bottom}: residual.  The data are binned from 2 to 10 minute
		cadence as a convenience for plotting and fitting.
		\label{fig:splitsignal}
	}
\end{figure*}

\begin{figure*}[hbtp]
	\begin{center}
		\leavevmode
		\includegraphics[width=1\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS lightcurve of \ptfo\ (Sector 6, Orbit 20).}
		Same format as Figure~\ref{fig:splitsignal}.
		\label{fig:splitsignalii}
	}
\end{figure*}

@@

\ptfo\ [also known as CVSO\,30; \cite{briceno_cida_2005}}]
was observed by TESS with Camera 1, CCD 1, from December 15,
2018 until January 6, 2019, during the sixth sector of science operations
\citep{ricker_transiting_2015}.  The star was designated TIC\,264461976
in the TESS Input Catalog \citep{stassun_TIC_2018,stassun_TIC8_2019}.
The pixel data for an $11\times11$ array surrounding \ptfo\ were
averaged into 2-minute stacks by the onboard computer.  Each
2048$\times$2048 image from the CCD was also averaged into 30-minute
stacks, and saved as a ``full frame image'' (FFI).

The 2-minute stacks for \ptfo\ were reduced to lightcurves by the
Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.  We mainly used the
Presearch Data Conditioning (PDC) lightcurve.  The PDC lightcurve
aperture used pixels chosen to maximize the SNR of the total flux of
the target \citep{smith_kepler_apertures_2017}.  Non-astrophysical
variability was removed by fitting out trends common to many stars
\citep{smith_kepler_PDC_2017}.

As an independent check on the 2-minute SPOC lightcurve, we
examined the lightcurve based upon 30-minute image stacks which was
produced as part of the Cluster
Difference Imaging Photometric Survey (CDIPS;
\citealt{bouma_cluster_2019}).  Our CDIPS lightcurve of choice used a
circular aperture with radius 1 pixel.
% JNW: What does that mean? You used fractional pixels in calculating the pixel sums?

To clean the data, we removed all points with non-zero quality flags,
which indicate known problems
\citep[{\it e.g.},][]{tess_data_product_description_2018}.  We also
masked out the data from the first and last 6 hours of each orbit, since there are
often systematic red noise in the photometry during those times.  Both the CDIPS and PDC
lightcurves showed a clear discontinuous jump in the last few days
of orbit 20, which seemed likely to be an instrumental systematic effect.  We
correspondingly masked out the data with timestamps ranging from BJD\,2458488.3 until
the end of the orbit.  The PDC lightcurve initially had 15{,}678 points.  The
quality-flag cut removed 854 points; masking the orbit edges removed an
additional 716 points; and removing the data from the final few days of orbit 20 removed an
additional 1079 points.  After cleaning, 83\% of the initial flux
measurements remained.

We normalized the lightcurve by dividing out the median flux, and then opted
to subtract 1.0 to set the median value to zero, which simplified
subsequent interpretation.  Many
of these and subsequent processing steps were performed using
\texttt{astrobase}~\citep{bhatti_astrobase_2018}.

@@

\subsubsection{Astrometric measurements}

Between July 25, 2014 and May 23, 2016, Gaia measured about 300
billion centroid positions of 1{.}6 billion stars
\citep{gaia_collaboration_gaia_2016,lindegren_gaiasoln_2018,gaia_collaboration_gaia_2018}.
In the Gaia second data release (DR2), these CCD observations were
used to determine positions, proper motions, and parallaxes of the
brighest 1{.}3 billion stars, including \ptfo\
\citep{lindegren_gaiasoln_2018}.  For \ptfo, there were 121 ``good'' observations, i.e., observations that were not strongly down-weighted in
the astrometric solution.  \ptfo\ was assigned the Gaia DR2 identifier
3222255959210123904.  It photometric brightness was measured using
selected bands ($G$, $Rp$, and $Bp$) of the Gaia Radial Velocity
Spectrometer \citep{cropper_gaia_2018,evans_gaia_2018}.  We accessed
the pipeline parameters for \ptfo\ using the Gaia
archive\footnote{\url{gea.esac.esa.int/archive/}}.

The majority of Gaia's derived parameters for \ptfo\ agreed with
expectations from former studies
\citep{briceno_cida_2005,van_eyken_ptf_2012}.  The main novelty is
that Gaia DR2 reported a 10.3$\sigma$ ``astrometric excess'', indicating
that the residuals to the best-fitting astrometric model were larger
than expected based on the statistical uncertainties.
We comment on the significance and
interpretation of this excess in Section~\ref{sec:gaia}.

@@

Gaia also provided astrometric parameters for tens of thousands of
young stars in the Orion complex.  Stellar populations in giant
molecular cloud complexes are not monolithic; substructured groups are
the norm \citep{briceno_lowmassOB_2007}.  The Orion molecular cloud
complex in particular has numerous subgroups, with ages ranging from 0.5
to 15$\,$Myr. For an incomplete sampling, see for instance
\citet{briceno_cida_2005,jeffries_kinematic_2006,briceno_25_2007,kounkel_apogee2_2018}
and \citet{briceno_cidaII_2019}.

\ptfo\ was initially identified as a member of the Orion$\,$OB1a
sub-association and designated CVSO\,30 by \citet{briceno_cida_2005} through combined
photometry and spectroscopy.  Later work by \citet{briceno_25_2007}
clarified that \ptfo\ was in a kinematically distinct subgroup of
Orion~OB1a, named the ``25$\,$Ori'' group after its brightest
member. They reported that the 25$\,$Ori group has an isochrone age of 7--10$\,$Myr, and
a smaller fraction of stars with disks than younger nearby sub-associations
\citep{hernandez_spitzer_ob1_2007}.

With the Gaia astrometry, it has become clear that 25$\,$Ori itself
has distinct subgroups
\citep{kounkel_apogee2_2018,briceno_cidaII_2019}.  In describing the
cluster membership of \ptfo, we follow the notation and results of
\citet{kounkel_apogee2_2018}.  These authors combined astrometric data
from Gaia DR2 with near-infrared spectra from APOGEE-2
\citep{gunn_sdss_2006,majewski_apache_2017,blanton_sloan_2017,zasowski_target_2017,cottle_apogee2_2018}.
They performed a hierarchical clustering on the six-dimensional
position and velocity information to identify subgroups within the
Orion complex.  From smallest to largest, \ptfo\ was identified
as a member of the following hierarchical subgroups:
\begin{equation}
  %{\rm PTFO\,8\text{-}8695}
  % \in
  {\rm 25\,Ori\text{-}1}
  \subset {\rm 25\,Ori}
  \subset {\rm Orion\ OB1a}
  \subset {\rm Orion\ D},
\end{equation}
where `$\subset$' means ``is a proper subset
of''.  25$\,$Ori-1 is the largest subgroup of 25$\,$Ori, with 149
identified members.  Based on the CMD, the mean age was found to be
6.9$\,$Myr, while the HR diagram led to an estimated age of 8.5$\,$Myr
\citep{kounkel_apogee2_2018}. 
% JNW: What's the difference between a CMD and an HR diagram?
\citet{kounkel_apogee2_2018} identified
seven other smaller groups in the Orion complex near the Be star
25$\,$Ori. These groups received higher numbers, {\it e.g.},
25$\,$Ori-2 (${\rm Age}_{\rm CMD} =15.1\,$Myr; ${\rm Age}_{\rm CMD}
=12.9\,$Myr; see also \citealt{briceno_cidaII_2019}).

These details concerning the group membership for one object are
cumbersome to those accustomed to the simple distinction between ``young cluster members''
and ``old field stars''.  Although all members of the Orion complex are
indeed young relative to the field, these details are essential for
assessing the photometric evidence for the binarity of \ptfo, because of
the degeneracy between stellar luminosity and age for
pre-main-sequence stars.  Having a clean sample of reference stars that are
tightly associated with \ptfo\ --- both spatially
and kinematically --- minimizes contamination
not only from field stars, but also from older and younger members of the
Orion complex.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f3.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf Phase-folded longer and shorter-period signals.}
    {\it Top}: Longer-period signal, as in Figure~\ref{fig:splitsignal}.
    {\it Bottom}: Shorter-period signal. The reference phase is set to
    the dip.  Gray points are the 10 minute cadence
    \texttt{PDCSAP} flux.  Black points are binned to 100 points per
    period.
		\label{fig:phasefold}
	}
\end{figure*}

@@

Our initial inspection of the TESS lightcurve, in both its 2-minute
PDCSAP and 30-minute FFI forms, showed a strong sinusoidal beat signal
(Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}, top
panel). As a precursor to more detailed analysis, we calculated generalized
Lomb-Scargle periodograms using \texttt{astrobase}
\citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015,bhatti_astrobase_2018}.
The two tallest peaks occur at 0.448\,days and 0.499\,days.
We will refer to these two periods as the ``shorter period'' $P_{\rm s}$
and the ``longer period'' $P_{\rm ell}$.
The peak at $P_{\rm \ell}$ is the taller of the two peaks.
Lower-power harmonics of both signals are also present.

@@

The peak-to-peak lightcurve amplitude at maximum, when the two signals
interfere constructively, is about 14\%.  During the times of destructive interference, the peak-to-peak
amplitude is about 6\%.  Assuming the signals are mainly sinusoidal,
simple algebra tells us that the peak-to-peak amplitudes should be
about 10\% for the longer-period signal, and 4\% for the shorter-period signal.
To view the phase-folded light curves of the longer-period signal,
we subtracted the best-fitting sinusoid at the shorter period; the resulting
light curve appears smooth and nearly sinusoidal.
But after subtracting the best-fitting sinusoid at the longer period,
visual inspection of the phase-folded light curve of the shorter-period signal
revealed substructure resembling the ``dips'' seen in previous
observations. In particular, there was a $\approx$1\% dip lasting about an hour.
These initial impressions turned out to be consistent with the results of our
more complicated analysis, described below.

% JNW This doesn't seem necessary to me.
%Initial signal-processing experiments fitting out splines or sinusoids
%showed that after subtracting out the long-period signal, the
%short-period signal dominated the periodogram, and vice-versa.
%However it quickly became clear that it would be beneficial to
%simultaneously model each signal, in order to preserve the
%power at each frequency.

@@

We considered nine models, with the number of modes per frequency
$N$ and $M$ ranging from one to three.  To select our preferred model,
we used the Bayesian information criterion (Table~1).  The model with
the lowest BIC had two modes at the shorter 10.74$\,$hr period, and
two modes at the longer 11.96$\,$hr period.  The next-best model had
an additional harmonic at the longer period ($M=3$), but was otherwise
identical.  All nine models have reduced $\chi^2$ values ranging between 1.37
and 1.51, which suggests a plausible though imperfect agreement
between the data and the model to within the formal uncertainties.
Table~2 gives the best-fitting parameters for the preferred model,
which has the lowest BIC value.

@@

\subsection{Longer-Period Signal}

The standard interpretation for 11.96$\,$hour nearly sinusoidal modulations
of a pre-main-sequence M dwarf is stellar rotation.  This is the
dominant signal in the system with 10\% amplitude, and there is no
evidence to suggest that this signal has any other origin.
